\documentclass[../../../OAE-SPEC-MAIN.tex]{subfiles}
\begin{document}

\section{Our Unfair Advantage in Networking}

Conventional Clos Networks assume a “flat” communication model, and an Any to Any (A2A) traffic pattern.  What if we could take advantage of spatial (neighborhood) and temporal (synchronization) traffic patterns and provide ultra low latency in directly connected (switchless) networks?

\subsection{ML/AI Inference on The Edge}

While training is for the big boys, and often needs gigantic capital and operating budgets, inference can be carried out with more modest budgets, on far smaller networks, fewer GPUs and CPUs and is highly constrained by latency requirements of customers whose attention span has been trained by expectations of search engine response times.

These latency characteristics are also naturally serviced from the edge where the customers are, rather than the big centralized datacenters owned by the superscalars.

This provides many more market insertion opportunities for DÆDÆLUS’ unique insights and unfair advantage as a software infrastructure company bringing new capabilities and value to the edge. 

\subsection{Local not Global Traffic Patterns}
Datacenter Traffic Matrices are sparse and skewed, exhibit locality and are bursty. [Griner SIGMETRICS 2020]. They are far from arbitrary or random, which is often used as an justification for a “Flat” networks model in datacenters.

Reconfigurable networks [Dynamic Trees]  allow to establish direct links between two frequently communicating pairs of racks in a datacenter, e.g., using digital micromirror devices.

\subsection{Snapshots need Reliable Subtransactions}
Both Transactions (MVCC) and AI/ML rely on snapshots being consistent. But they can’t be made consisted using timestamps alone.

\subsection{Spanning Trees under application control}

\subsection{Failure, isolation, healing: using local information only}
\subsection{Topology Awareness for Applications like AI/ML}

\subsection{Sequencing and clocks without timestamps}

Timestamps are an unsafe ordering construct.

\subsection{Our unfair advantage}
\begin{itemize}
  \item Eliminate timeouts \& retries, which cause cascade failures
  \item Switches and routers not needed
  \item SmartNICs or Chiplet microdatacenters
  \item Easy to use (compared to conventional networks)
  \item Links carry only functional (not transit) packets
  \item More potential Bandwidth at the same cost
  \item Ultra Low Latency is lowest possible: governed only by the length of the connections
  \item Connections - Ours are are shorter.  (2cm vs 2m vs 20m)
  \item Encryption not needed between adjacent nodes (Saves computation, time, and energy)
  \item Developers control their own application topologies
  \item No bandwidth amplification, either by Gossip protocols or timeouts and retries
\end{itemize}



\subsection{Opportunities we Unleash}

HammingMesh is Optimized specifically for deep learning workloads and their communication patterns.

\begin{quote}
“Deep learning training uses three dimensional communication patterns and rarely needs global bandwidth. It supports extreme local bandwidth while controlling the cost of global bandwidth."
\end{quote}

HammingMesh uses a 4x4 Tile, connected in a valency-4 (square) mesh, and overlays a Hamming code to provide a GEV coordinate system for routing.  As with many coordinate systems, including the cartesian coordinate systems used in most Torus networks, they become extremely complex when they fail and we have to route around failed links and nodes.  This is not the case for DÆDÆLUS, where every node has an  coordinate system with itself at the root. Because 

The Strawman objection, by conventional network architects, uses “diameter” as a figure of merit (or demerit). Includes three misconceptions:
\begin{itemize}
  \item Nodes are “fixed” in some Cartesian layout, nailed to the floor of the datacenter in some specific rack (our addressable entities are mobile)
  \item Hops are somehow related to latency in a fixed cartesian coordinate system, so the diameter gets worse as you scale (DÆDÆLUS exploits neighborhood locality)
  \item Execution entities are randomly placed throughout the datacenter — for a “flat network concept”. %(DÆDÆLUS’ deployments are naturally clustered on deployment, AND can be migrated toward each other with a spring-electrical embedding in the orchestration protocols).
  \item Most AI/ML Traffic patterns are highly predictable. Once the pattern is established in the first few passes, the communication patterns can be exploited and the virtual trees can be optimized to match those patterns.
  %\item We're an early stage Silicon Valley company building software infrastructure for distributed systems that provides the illusion of an unbreakable network. We do that to make transactions reliable. We came out of stealth in June this year with our hypertransaction protocol.
\end{itemize}


\end{document}
